---
title: "Getting Started with raveR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with raveR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

raveR transforms R scripts into 'deep house' music. It analyzes your code
structure --- functions, control flow, complexity --- and maps it to musical
parameters, producing a deterministic audio track. The same code always
generates the same music.

## Installation

```{r}
# Install from CRAN
install.packages("raveR")

# Or development version from GitHub
devtools::install_github("voltalino/raveR")
```

MP3 export requires FFmpeg. WAV export works out of the box.

## Basic usage

The main function is `composeR()`. Point it at any R script and an output
path:

```{r}
library(raveR)

composeR("my_analysis.R", "my_analysis.wav")
```

That's it. raveR parses the script, extracts structural features, maps them
to musical parameters, and writes a complete 'deep house' track.

### Choosing a tempo

By default, raveR picks a tempo (118--124 BPM) deterministically from the
file hash. You can override it:

```{r}
composeR("script.R", "track.wav", bpm = 122)
```

### Reproducibility

Output is deterministic by default --- the same script always produces the same
audio. You can also set an explicit seed:

```{r}
composeR("script.R", "track.wav", seed = 42)
```

### MP3 export

If FFmpeg is installed, you can export directly to MP3:

```{r}
composeR("script.R", "track.mp3")
```

## Live mode

`playR()` starts background playback that responds to code changes in
real-time. Edit and save your script, and the music transitions smoothly:

```{r}
playR("my_analysis.R")

# Edit and save my_analysis.R --- music transitions
# ...keep working...

stopR()
```

The R console stays interactive during playback. Live mode requires the
`watcher` package (`install.packages("watcher")`).

## How the mapping works

raveR extracts structural features from your code and maps them to musical
parameters:

| Code feature            | Musical parameter       |
|-------------------------|-------------------------|
| Function count          | Number of musical parts |
| Nesting depth           | Harmonic complexity     |
| Control flow statements | Pattern density         |
| Variable count          | Instrument layers       |
| Cyclomatic complexity   | Arrangement length      |

A script with many nested loops and complex control flow produces a denser,
longer track with more harmonic variation. A simple script with one function
and a few assignments produces a minimal, stripped-back piece.

### Arrangement structure

Every track follows a standard 'deep house' arrangement:

1. **Intro** --- drums fade in, establishing the groove
2. **Build** --- bass and pads layer in, tension rises
3. **Drop** --- all elements playing, full energy
4. **Breakdown** --- stripped back, pads and atmosphere
5. **Outro** --- elements fade out

The number of bars in each section is determined by your code's complexity.

### Musical elements

- **Kick**: four-on-the-floor pattern (the defining 'deep house' rhythm)
- **Hi-hats**: 16th notes with swing for groove
- **Snare/clap**: beats 2 and 4
- **Bass**: filtered sawtooth, pattern derived from code structure
- **Pads**: warm chord progressions (min7, maj7, min9 voicings)
- **Lead**: melodic motifs generated from function names

### Determinism

The mapping uses MD5 hashing throughout. Function names hash to specific
musical parts, variable names hash to instruments, and the file path
determines the key and chord progression. This means:

- Same script, same music (always)
- Renaming a function changes its musical role
- Adding a variable adds an instrument layer
- Increasing nesting depth deepens the harmony

## Genres

raveR supports several genre presets that adjust tempo ranges and synthesis
parameters:

```{r}
composeR("script.R", "track.wav", genre = "deep_house")  # default
composeR("script.R", "track.wav", genre = "techno")
composeR("script.R", "track.wav", genre = "ambient")
composeR("script.R", "track.wav", genre = "drum_bass")
```
